Human Factors Influencing Visual Statistical Inference

### 2015-05-19 sociology_paper.tex
checked through changes completely

\begin{table*}[hbtp] 
\centering 
\caption{Overview of 10 different Turk experiments, \hh{in which} data was \hh{collected} to study human factor effects. All of the experimental data was used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.} 
\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
\hline\hline 
 & Experiment &  Test Statistic  & Lineup question & Used in study of\\ [0.5ex] % inserts table %heading 
\hline 
1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
between group A and B? & DF\\
2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope? & DF\\
  3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope? & DF\\
 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\  
  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group? &  DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups? & DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\ 
  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent? & DF\\ 
  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small? & DF  \hfill \phantom{LT} \hfill {LE} \hfill \phantom{.}\\ 
  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others? & DF\\

@@

We have conducted a series of experiments (see Table \ref{tbl:visual_stat} for an overview)
%There have been ten experiments conducted 
using Amazon's Mechanical Turk~\citep{turk} (MTurk) that provide a basis for evaluating
%are being used to evaluate 
the effects of different covariates on an individual's ability to evaluate lineups. 
In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.  
 %Each of these experiments collected demographic details of the subjects. 
 Experiments~5, 6 and 7 can be used to investigate short term learning trends. Experiment~9 was also designed to check whether the  positioning of the  data plot in the lineup had an effect on detection rate.

 Section \ref{sec:factor_performance} discusses human factors that may affect the performance of the observer. Section \ref{sec:exp_design} describes the methods used to assess the effects, and Section \ref{sec:result_socio} describes the results.

@@

\hh{Visual statistical inference relies not only on the strength of the signal in the data, but also on how this information is presented in the plot, on the lineup design and on the abilities of human observers.} It is important to understand how these factors might affect results.  \hh{Here, we provide a brief discussion }%This section provides a brief description 
of the factors that are expected to have some impact on the performance of visual statistical inference.

@@

\begin{itemize} \itemsep 0in \parsep-10pt %\parskip0pt 
%\subsection{Choice of Visual Test Statistic}  
\item{\bf Choice of Visual Test Statistic:}  
%There are a variety of ways that
Data can be plotted in a variety of ways and we can enhance plots by axes, legends or representations of models. The choice of plot primarily should be appropriate for the problem being investigated, as the data plot is the visual test statistic. 
When for example studying the association between two quantitative variables a scatterplot is usually \hh{considered best practice}. Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.  Some choices of visual test statistics provide better results than others when studying the same problem \citep{heike:2012}, because some plot types make it easier for certain structures to be seen.

@@

%\subsection{Signal in the Data}
\item{\bf Signal in the Data:}
A visual test statistic is chosen for a particular task. If it is well-designed then departures from the null hypothesis will be visible in the plot, especially if the signal is strong. An important factor that helps an observer to identify the actual data plot in a lineup is the strength of the signal in the data. Simulated data was generated for all of the experiments, except exps.~4 and~9, enabling the study of signal strength and actual data plot detection. \hh{In experiment~4 we took a different approach by varying the sampling size, which also allowed us to investigate the effect a design had on detection rate while varying signal/noise ratio.} Primarily the results were as expected, that as the signal strength increased observers more frequently picked the actual data plot as different from the null plots.

@@

In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted. This was important \hh{to allow a direct comparison of the results to results from conventional testing. }%in order to make a comparison of the results. 
For latter studies (4, 7, 10), questions posed to the observer were more general. Table \ref{tbl:visual_stat} lists the questions used for each experiment.

@@

%\subsection{Demographics} 
\item{\bf Demographics:}  \hh{During each of the experiments } data on age, gender, education level and geographic location \hh{was collected}. Each observer self-reported gender, age in roughly five year intervals, education level as high school or less, some college courses, an undergraduate degree, some graduate courses or a graduate degree. The IP address of the computer afforded the geographical location of the subject. The purpose of collecting this information with each experiment is to examine the effect that they have on the results of visual inference -- ideally very little.

@@

%\subsection{Location of Actual Data Plot in the Lineup}
\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 \hh{panels} is used. A random number generator is used to determine the position where the actual data plot is placed in each lineup. Eye tracking experiments \citep{zhao:2012} suggest that some observers traverse lineups in horizontal direction, while others have a vertical up-and-down approach to their search. Almost universally, observers start at the top left of a lineup. This leads to some concern that observers might be able to more easily identify the actual data plot, if it is placed at the top left of a lineup than when it the actual data is placed at the bottom right.
%With the help of eye-tracking equipment \citet{zhao:2012} tracked the observers' eyes to see their approach in searching the plots in a lineup. The results suggest that many people read lineups from left to right, some read from upward to downward. 
%There is some concern that if the actual data plot is located in the top left of the lineup, observers may be able to identify it more easily than when the actual data plot is at the bottom right. 
Ideally, this does not happen. Experiment~9  is designed to allow us to investigate the effects of location on detection. Five different locations in the lineup were used to assess how fast and accurately observers identified the actual data plot.

@@

%\subsection{Individual Skill or Ability} 
\item{\bf Individual Skill or Ability:} Each person may have different  aptitude for reading statistical plots, and their visual skill sets might be differently developed.  \hh{We can evaluate the effect of individuals' skills and abilities}  by letting multiple subjects evaluate the same lineups, which allows us to estimate, if some subjects consistently detect the actual data plot more often than others.  
 %For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision. This influences the response of the observer. Also subject specific variation in the power of visual test is observed in \cite{majumder:2013}. 
\end{itemize}

@@

\subsection{Overview of the Data} A total of 2321 participants provided feedback data on the lineups in ten different experimental studies. Figure~\ref{fig:turker_location} displays the locations of participants around the world.  Most of the participants were from the United States and India. There were 76 other different countries represented. This provides a diverse pool of participants. The diversity in not only  geographic but also in gender, age group and education level as can be seen in Table \ref{tbl:demographics}. % It is interesting that there were large number of female participants even though there were a lot of participants  from developing countries.
\hh{The large number of female participants from all countries was a pleasant surprise to us.}

@@

The largest number of participants \hh{falls within the} age group of 18 to 25, with the majority between 18 to 35. \hh{ Many older participants took part in the experiments.} %Interestingly, there are many participants in the older age groups as well. 
The United States shows a more uniform participation beyond age 30, Indian participants are primarily younger.

In terms of academic background, \hh{a majority of participants has a graduate degree} (a total of 902 participants, about 38.51\%). Examining the location in conjunction with the academic background reveals that there are more Indian participants that \hh{report to have } a graduate degree. In retrospect \hh{we realized this to be a culturally based misunderstanding}, as any university degree in India is considered to be a graduate degree, while the North American definition of graduate degree refers to a Masters level education. Most of the USA participants \hh{report to have} an undergraduate degree or at least have some undergraduate courses.

The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where \hh{the} proportion of female participants is lower. The distribution of education levels also differs slightly across countries for this age group. %Most of the participants from India are below age 40 while in United States the distribution of participants are similar beyond age 40. There were not many participants beyond age 50 and hence these age groups are merged to form one age group called above 50. The exploratory analysis and the models fitted in the following sections use this new age group instead of actual age groups beyond age 50.

A total of 1911 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where three lineups were evaluated by each person. 
\hh{In order to ensure data quality, a test plot was shown to each observer, and the performance on this test plot was used for screening purposes.} %A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality. 
\hh{Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.}
Data from participants who \hh{gave the wrong answer on the test plot were considered as gaming the system, and their data} were removed.
\hh{Results from test plots were not considered in any of the analyses}.
 In some cases the demographic information were not provided by the participants. Also, for some ip address, the actual geographical locations could not be retrieved. This resulted in some missing demographic information.

@@

Parameter estimates from the model fits are shown in Table \ref{tbl:model_result_demographics}. The first level of each factor serves as the baseline for the model. The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for age groups 31-35 and above 50. Participants from India have a similar detection rate to those from the USA, but \hh{participants} from the rest of the world show a significantly \hh{higher} detection rate. Subjects  \hh{reporting to have}  a graduate degree had a significantly higher detection rate. For time to respond, all age groups were significantly slower than  18-25 year olds.  Participants from India were slower, while those from the rest of the world were faster than US Americans. A higher degree corresponded to longer time, and males were slower than females.

@@

Table \ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time} where response time is examined with respect to attempt. Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated. Interestingly, time to respond significantly decreases as number of attempts increases, for all three experiments. 
The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments. The negative estimates suggest that on average later attempts took less time. Even though observers did not improve in their performance in successive attempts, they became more efficient in their response. The parameter $\alpha_1$ for first attempt is also highly significant. The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts. \hh{A possible reason is that an observer might have gone through instructions and familiarized themselves  with the experimental environment.} Each of the experiments \hh{asks observers to give a reason for their choice (out of a list of pre-defined answers}. \hh{For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).}  Later pages of the web site were identical  except for the lineup. Thus in the later attempts an observer does not need to spend any time for reading instructions. The model reflects that fact.

@@

We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets. The lineup protocol is using a finite and small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set. If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate. From the plot we can see that null set 3 appeared to be more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot. This null set effect was tested, and a significant difference \hh{was found} for null set 3 of the interaction lineup. The rest of the null sets do not show any differences at  a 1\% significance level.

@@

For the second approach in investigating the location effect, we distinguish between inner panels 7,8,9,12,13,14 and outer panels (any panel with an outside edge).  
%We also examined whether the proportion of correct responses differs if the actual plot is on the outer boundary or in the inner locations. The locations 7,8,9,12,13,14 in the lineup are considered inner locations and the rest of the boundary locations are considered to be outer location. 
From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype. 
Model \eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.  
%It does not show any differences whether the actual plot is inside or outer border of the lineup.

>>>>

\begin{table*}[hbtp] 
\centering 
\caption{Overview of 10 different Turk experiments, from which data was collected to study human factor effects. All of the experimental data was used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.} 
\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
\hline\hline 
 & Experiment &  Test Statistic  & Lineup question & Used in study of\\ [0.5ex] % inserts table %heading 
\hline 
1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
between group A and B? & DF\\
2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope? & DF\\
  3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope? & DF\\
 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\  
  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group? &  DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups? & DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\ 
  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent? & DF\\ 
  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small? & DF  \hfill \phantom{LT} \hfill {LE} \hfill \phantom{.}\\ 
  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others? & DF\\

@@

We have conducted a series of experiments (see Table \ref{tbl:visual_stat} for an overview)
%There have been ten experiments conducted 
using Amazon's Mechanical Turk~\citep{turk} (MTurk) for a variety to purposes: validating the protocol against existing tests, comparing plot designs, and evaluating structure in data analysis problems. In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.  Mining this data will provide a basis for evaluating
%are being used to evaluate 
the effects of different covariates on an individual's ability to evaluate lineups.  Experiments~5, 6 and 7 can be used to investigate short term learning trends. Experiment~9 was also designed to check whether the  positioning of the  data plot in the lineup had an effect on detection rate. 
 %Each of these experiments collected demographic details of the subjects. 
Section \ref{sec:factor_performance} discusses human factors that we expect may affect the performance of the observer. Section \ref{sec:exp_design} describes the methods used to assess the effects, and Section \ref{sec:result_socio} describes the results on what we have learned about human factors affecting lineup evaluation.

@@

Visual statistical inference relies not only on the strength of the signal in the data, but also on how this information is presented in the plot, on the lineup design and on the abilities of human observers. It is important to understand how these factors might affect results.  Here, we provide a brief discussion %This section provides a brief description 
of the factors that are expected to have some impact on the performance of visual statistical inference.

@@

\begin{itemize} \itemsep 0in \parsep-10pt %\parskip0pt 
%\subsection{Choice of Visual Test Statistic}  
\item{\bf Choice of Visual Test Statistic:}  
%There are a variety of ways that
Data can be plotted in a variety of ways and we can enhance plots by axes, legends or representations of models. The choice of plot primarily should be appropriate for the problem being investigated, as the data plot is the visual test statistic. 
When for example studying the association between two quantitative variables a scatterplot is usually considered best practice. Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.  Some choices of visual test statistics provide better results than others when studying the same problem \citep{heike:2012}, because some plot types make it easier for certain structures to be seen.

@@

%\subsection{Signal in the Data}
\item{\bf Signal in the Data:}
A visual test statistic is chosen for a particular task. If it is well-designed then departures from the null hypothesis will be visible in the plot, especially if the signal is strong. An important factor that helps an observer to identify the actual data plot in a lineup is the strength of the signal in the data. Simulated data was generated for all of the experiments, except exps.~4 and~9, enabling the study of signal strength and actual data plot detection. In experiment~4 we took a different approach by varying the sampling size, which also allowed us to investigate the effect a design had on detection rate while varying signal/noise ratio. Primarily the results were as expected, that as the signal strength increased observers more frequently picked the actual data plot as different from the null plots.

@@

In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted. This was important to allow a direct comparison of the results to those from conventional testing. %in order to make a comparison of the results. 
For latter studies (4, 7, 10), questions posed to the observer were more general. Table \ref{tbl:visual_stat} lists the questions used for each experiment.

@@

%\subsection{Demographics} 
\item{\bf Demographics:}  During each of the experiments, data on age, gender, education level and geographic location was collected. Each observer self-reported gender, age in roughly five year intervals, education level as high school or less, some college courses, an undergraduate degree, some graduate courses or a graduate degree. The IP address of the computer afforded the geographical location of the subject. The purpose of collecting this information with each experiment is to examine the effect that they have on the results of visual inference -- ideally very little.

@@

%\subsection{Location of Actual Data Plot in the Lineup}
\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 panels is used. A random number generator is used to determine the position where the actual data plot is placed in each lineup. Eye tracking experiments \citep{zhao:2012} suggest that some observers traverse lineups in horizontal direction, while others have a vertical up-and-down approach to their search. Almost universally, observers start at the top left of a lineup. This leads to some concern that observers might be able to more easily identify the actual data plot, if it is placed at the top left of a lineup than when it the actual data is placed at the bottom right.
%With the help of eye-tracking equipment \citet{zhao:2012} tracked the observers' eyes to see their approach in searching the plots in a lineup. The results suggest that many people read lineups from left to right, some read from upward to downward. 
%There is some concern that if the actual data plot is located in the top left of the lineup, observers may be able to identify it more easily than when the actual data plot is at the bottom right. 
Ideally, this does not happen. Experiment~9  is designed to allow us to investigate the effects of location on detection. Five different locations in the lineup were used to assess how fast and accurately observers identified the actual data plot.

@@

%\subsection{Individual Skill or Ability} 
\item{\bf Individual Skill or Ability:} Each person may have different  aptitude for reading statistical plots, and their visual skill sets might be differently developed.  We can examine the effect of individuals' skills and abilities because multiple subjects evaluated the same lineups, which allows us to estimate, if some subjects consistently detect the actual data plot more often than others.  
 %For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision. This influences the response of the observer. Also subject specific variation in the power of visual test is observed in \cite{majumder:2013}. 
\end{itemize}

@@

\subsection{Overview of the Data} A total of 2321 participants provided feedback data on the lineups in ten different experimental studies. Figure~\ref{fig:turker_location} displays the locations of participants around the world.  Most of the participants were from the United States and India. There were 76 other different countries represented. This provides a diverse pool of participants. The diversity in not only  geographic but also in gender, age group and education level as can be seen in Table \ref{tbl:demographics}. % It is interesting that there were large number of female participants even though there were a lot of participants  from developing countries.
The large number of female participants from all countries was a pleasant surprise to us.

@@

The largest number of participants falls within the age group of 18 to 25, with the majority between 18 to 35. Many older participants also took part in the experiments.%Interestingly, there are many participants in the older age groups as well. 
The United States shows a more uniform participation beyond age 30, Indian participants are primarily younger.

In terms of academic background, the majority of participants had a graduate degree (a total of 902 participants, about 38.51\%). Examining the participants' location in conjunction with the academic background reveals that there are more Indian participants that report to have a graduate degree. In retrospect we realized this to be a culturally based misunderstanding, as any university degree in India is considered to be a graduate degree, while the North American the definition of graduate degree refers to a Masters level, or above, education. Most of the USA participants report to have an undergraduate degree or at least have some undergraduate courses.

The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where the proportion of female participants is lower. The distribution of education levels also differs slightly across countries for this age group. %Most of the participants from India are below age 40 while in United States the distribution of participants are similar beyond age 40. There were not many participants beyond age 50 and hence these age groups are merged to form one age group called above 50. The exploratory analysis and the models fitted in the following sections use this new age group instead of actual age groups beyond age 50.

A total of 1911 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where three lineups were evaluated by each person. 
In each of the experiments, to ensure quality evaluations, a test plot was shown to each observer, and the performance on this test plot was used for screening purposes.%A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality. 
Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.
Data from participants who gave the wrong answer on the test plot were considered as gaming the system, and their data were removed.
Results from test plots were not considered in any of the analyses.
 In some cases the demographic information were not provided by the participants. Also, for some ip address, the actual geographical locations could not be retrieved. This resulted in some missing demographic information.

@@

Parameter estimates from the model fits are shown in Table \ref{tbl:model_result_demographics}. The first level of each factor serves as the baseline for the model. The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for age groups 31-35 and above 50. Participants from India have a similar detection rate to those from the USA, but participants from the rest of the world show a significantly higher detection rate. Subjects  reporting to have  a graduate degree had a significantly higher detection rate. For time to respond, all age groups were significantly slower than  18-25 year olds.  Participants from India were slower, while those from the rest of the world were faster than US Americans. A higher degree corresponded to longer time, and males were slower than females.

@@

Table \ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time} where response time is examined with respect to attempt. Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated. Interestingly, time to respond significantly decreases as number of attempts increases, for all three experiments. 
The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments. The negative estimates suggest that on average later attempts took less time. Even though observers did not improve in their performance in successive attempts, they became more efficient in their response. The parameter $\alpha_1$ for first attempt is also highly significant. The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts. Most observers, before evaluating the first lineup, carefully read through instructions and familiarize themselves  with the types of plots used in the experiment, which would inflate the time taken on the first lineup. Each of the experiments asks observers to give a reason for their choice (out of a list of pre-defined answers. For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).  Later pages of the web site were identical  except for the lineup. Thus in the later attempts an observer does not need to spend any time for reading instructions. The model reflects that fact.

@@

We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets. The lineup protocol is using a finite and small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set. If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate. From the plot we can see that null set 3 appeared to be more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot. This null set effect was tested, and a significant difference was found for null set 3 of the interaction lineup. The rest of the null sets do not show any differences at  a 1\% significance level.

@@

For the second approach in investigating the location effect, we distinguish between inner panels 7, 8, 9, 12, 13, 14 and outer panels (any panel with an outside edge).  
%We also examined whether the proportion of correct responses differs if the actual plot is on the outer boundary or in the inner locations. The locations 7,8,9,12,13,14 in the lineup are considered inner locations and the rest of the boundary locations are considered to be outer location. 
From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype. 
Model \eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.  
%It does not show any differences whether the actual plot is inside or outer border of the lineup.
